---
title: "Housing Data"
author: "Melissa Young"
date: "July 28, 2022"
output: pdf_document
---

```{r, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/mdyoung/OneDrive - Bankers Financial Corporation/Documents/R/DSC 520' )
library(plyr)
library(dplyr)
library(lmtest)
library(readxl)
library(tidyverse)
library(car)
library(lubridate)
library(outliers)
library(ggplot2)
```


Housing Data

Work individually on this assignment. You are encouraged to collaborate on ideas and strategies pertinent to this assignment. Data for this assignment is focused on real estate transactions recorded from 1964 to 2016 and can be found in Housing.xlsx. Using your skills in statistical correlation, multiple regression, and R programming, you are interested in the following variables: Sale Price and several other possible predictors.

If you worked with the Housing dataset in previous week - you are in luck, you likely have already found any issues in the dataset and made the necessary transformations. If not, you will want to take some time looking at the data with all your new skills and identifying if you have any clean up that needs to happen.


```{r, echo=FALSE}
df <- read_excel("C:\\Users\\mdyoung\\OneDrive - Bankers Financial Corporation\\Documents\\R\\DSC 520\\week-6-housing.xlsx")
```




**Complete the following:**

**1. Explain any transformations or modifications you made to the dataset**

a. Separated the Sale Date into coulumns for year, month and day. 
```{r, include=FALSE}
df %>%
mutate(
        year = substr(df$`Sale Date`,  1, 4),
        month = substr(df$`Sale Date`,  6, 7),
        day = substr(df$`Sale Date`, 9, 10))
```

b. Added a calculation for Price Per Sqft (Price per square foot) and Age of Home in Year Sold. 
```{r, include=FALSE}
df %>%
  mutate(`Price Per SqFt` = round(df$`Sale Price`/ df$square_feet_total_living,2),
         `Age of Home` = (df$year - df$year_built))
```

c. Added the state with a zip code lookup.
 

**2. Create two variables; one that will contain the variables Sale Price and Square Foot of Lot (same variables used from previous assignment on simple regression) and one that will contain Sale Price and several additional predictors of your choice. Explain the basis for your additional predictor selections.**

```{r}
single_lm <- lm(`Sale Price` ~ sq_ft_lot, data = df)

multi_lm <- lm(`Sale Price` ~ sq_ft_lot + sale_reason + building_grade + square_feet_total_living + bedrooms + bath_full_count + bath_half_count + current_zoning + present_use + `year` +`month` + `day` + `Price Per SqFt` + `Age of Home in Year Sold`, data = df)

```

**3. Execute a summary() function on two variables defined in the previous step to compare the model results. What are the R2 and Adjusted R2 statistics? Explain what these results tell you about the overall model. Did the inclusion of the additional predictors help explain any large variations found in Sale Price?**
```{r}
summary(single_lm)
summary(multi_lm)
```
a. Single Multiple R-squared:  0.01435,	Adjusted R-squared:  0.01428
b. Multi Multiple R-squared:  0.7309,	Adjusted R-squared:  0.7301
c. When R-squared and Adjusted R-Squared are closer to 1, the more accurate your model. From a and b above, the multi is more accurate than the single. 

**4. Considering the parameters of the multiple regression model you have created. What are the standardized betas for each parameter and what do the values indicate?**
```{r}
lm_betas <- data.frame(multi_lm$coefficients)
```
The values indicate that the variables impact the model with the degree of error it contains.  


**5. Calculate the confidence intervals for the parameters in your model and explain what the results indicate.**
```{r}
confidence_interval <- confint(multi_lm, level = 0.95)
```
The confidence interval results indicate that you can be 95% certain that values fall within that range of the true mean. 

**6. Assess the improvement of the new model compared to your original model (simple regression model) by testing whether this change is significant by performing an analysis of variance.**
```{r}
Anova(single_lm, multi_lm)
```


**7. Perform casewise diagnostics to identify outliers and/or influential cases, storing each function's output in a dataframe assigned to a unique variable name.**
```{r}
housingdata <- df

housingdata$residuals <- resid(multi_lm)
housingdata$standardized.residuals <- rstandard(multi_lm)
```


**8. Calculate the standardized residuals using the appropriate command, specifying those that are +-2, storing the results of large residuals in a variable you create.**
```{r}
housingdata$standardized.res <- housingdata$standardized.residuals >2 | housingdata$standardized.residuals < -2
```


**9. Use the appropriate function to show the sum of large residuals.**
```{r}
large_residuals <- housingdata %>%
  group_by(standardized.res) %>%
  summarise(standardized.residuals = sum(standardized.residuals)) %>%
  filter(standardized.res == T)
```


**10. Which specific variables have large residuals (only cases that evaluate as TRUE)?**
```{r}
large_residuals_variables <- housingdata %>%
  group_by(standardized.res) %>%
  filter(standardized.res ==T)
```


**11. Investigate further by calculating the leverage, cooks distance, and covariance rations. Comment on all cases that are problematics.**
```{r}
housingdata$cooks.distance <- cooks.distance(multi_lm)
housingdata$dfbeta <- dfbeta(multi_lm)
housingdata$dffit <- dffits(multi_lm)
housingdata$leverage <- hatvalues(multi_lm)
housingdata$covaraince.ratios <- covratio(multi_lm)
```


**12. Perform the necessary calculations to assess the assumption of independence and state if the condition is met or not.**
```{r}
independence_assumption <- dwt(multi_lm)
```


**13. Perform the necessary calculations to assess the assumption of no multicollinearity and state if the condition is met or not.**
```{r}
multicollinearity_assumption <- vif(multi_lm)
```


**14. Visually check the assumptions related to the residuals using the plot() and hist() functions. Summarize what each graph is informing you of and if any anomalies are present.**
```{r, warning=FALSE, message=FALSE}
plot(multi_lm)

ggplot(housingdata, aes(x = `dffit`, y = `standardized.residuals`)) + geom_point() + xlab("Values") + ylab("Residuals") + ggtitle("Values / Standardized Residuals")

ggplot(housingdata, aes(x = `standardized.residuals`)) + geom_histogram(binwidth = .50, color = "orange", fill = "green") + xlab("Residuals") + ggtitle("Residuals Histogram")
```


**15. Overall, is this regression model unbiased? If an unbiased regression model, what does this tell us about the sample vs. the entire population model?**

Overall the regression model is unbiased because it shows that the sampling distribution is equal to the true population parameter value. From the histogram above, would show a normal curve.  