---
title: "Housing_Dataset_YoungMelissa"
author: "Melissa Young"
date: "July 3, 2022"
output: pdf_document
---

```{r, echo = F}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```
# Load appropriate packages
library(tidyr)
library(pastecs)
library(plyr)
library(dplyr)
library(ggplot2)
library(readxl)
```{r,warning=FALSE,message=FALSE}
library(tidyr)
library(pastecs)
library(plyr)
library(dplyr)
library(ggplot2)
library(readxl)
library(formatR)
library(reshape)
```

```{r}
read_excel("C:\\Users\\mdyoung\\OneDrive - Bankers Financial Corporation\\Documents\\R\\DSC 520\\week-6-housing.xlsx")
```

```{r}
df <- read_excel("C:\\Users\\mdyoung\\OneDrive - Bankers Financial Corporation\\Documents\\R\\DSC 520\\week-6-housing.xlsx")
```



#1. Use the apply function on a variable in your dataset.
```{r}
apply(select(df, c(2,14)), 2, sum)
```
#2. Use the aggregate function on a variable in your dataset
```{r}
aggregate(df$`Sale Price` ~ df$'type', df, mean)
```
#3. Use the plyr function on a variable in your dataset - more specifically, I want to see you split some data, perform a modification to the data, and then bring it back together.
```{r}
df %>%
mutate(
        year = substr(df$`Sale Date`,  1, 4),
        month = substr(df$`Sale Date`,  6, 7),
        day = substr(df$`Sale Date`, 9, 10))
```

```{r}
by_year <-
  df %>%
  group_by(year) %>%
  summarise(`Total Sales by Year` = sum(`Sale Price`))
  
```
```{r}
by_year
```


#4. Check distributions of the data.
```{r}
df %>%
  group_by(year) %>%
  summarise(`Sale Price` = sum(`Sale Price`)/1000000) %>%
  ggplot(aes(x = year, y = `Sale Price`)) + 
  geom_bar(
    stat = 'identity',
    size = .5,
    color = 'black',
    fill = 'orange',
    alpha = 1) +
  ggtitle("Density check of Total Sales by Year")
```

#4. Check distributions of the data.
```{r}
by_year %>%
ggplot(aes(x = 'Total Sales by Year')) + geom_density(fill = 'orange') +
  stat_function(fun = dnorm, args = list(
    mean = mean(by_year$`Total Sales by Year`),
    sd = sd(by_year$`Total Sales by Year`))) + ggtitle("Density of Sales by Year")
```

#5. Identify if there are any outliers.
```{r}
boxplot((by_year$`Total Sales by Year`)/1000000)
boxplot.stats((by_year$`Total Sales by Year`)/1000000)$out

```
#6. Create at least 2 new variables.
```{r}
df %>%
  mutate(`Price Per SqFt` = round(df$`Sale Price`/ df$square_feet_total_living,2),
         `Age of Home` = (df$year - df$year_built))
```


