---
title: "Week_5_Transformations_YoungMelissa"
author: "Melissa Young"
date: "July 6, 2022"
output: pdf_document
---

```{r, echo = F}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

```{r,warning=FALSE,message=FALSE}
library(tidyr)
library(pastecs)
library(plyr)
library(dplyr)
library(ggplot2)
library(readxl)
library(purrr)
library(gapminder)
```

#1. Using either the same dataset(s) you used in the previous weeks' exercise or a brand-new dataset of your choosing, perform the following transformations (Remember, anything you learn about the Housing dataset in these two weeks can be used for a later exercise!)

```{r}
read_excel("C:\\Users\\mdyoung\\OneDrive - Bankers Financial Corporation\\Documents\\R\\DSC 520\\week-6-housing.xlsx")
```

```{r}
df <- read_excel("C:\\Users\\mdyoung\\OneDrive - Bankers Financial Corporation\\Documents\\R\\DSC 520\\week-6-housing.xlsx")
```

##1.  Using the dplyr package, use the 6 different operations to analyze/transform the data - GroupBy, Summarize, Mutate, Filter, Select, and Arrange - Remember this isn't just modifying data, you are learning about your data also - so play around and start to understand your dataset in more detai. 

```{r}
by_year <-
  df %>%
  group_by(year) %>%
  summarise(`Total Sales by Year` = sum(`Sale Price`))
  
```
```{r}
by_year
```
```{r}
df %>%
mutate(
        year = substr(df$`Sale Date`,  1, 4),
        month = substr(df$`Sale Date`,  6, 7),
        day = substr(df$`Sale Date`, 9, 10))
```
```{r}
df %>%
filter(`Sale Price`>4000000)
```
```{r}
df %>% select(square_feet_total_living, `Sale Price`) %>%
arrange(desc(`square_feet_total_living`))

```



##2. Using the purrr package - perform 2 functions on your dataset.  You could use zip_n, keep, discard, compact, etc.

```{r}
timesFive <- function(.x) {
  return(.x * 5)}
map(.x = by_year$`Total Sales by Year`, 
    .f = timesFive)
timesTwo <- function(.x) {return(.x*2)}

modify_if(.x = by_year$`Total Sales by Year`, .p = function(x) x > 500000000, .f = timesTwo)
```

##3. Use the cbind and rbind function on your dataset.
```{r}
Year <- c(2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013)
Sales <- c(919005546, 850285247, 755045696, 454417263, 592828305, 693256612, 747585171, 860712529)
df_2 <- cbind(Year, Sales)
df_2

```

```{r}
df_3 <- data.frame(Year=c(2014,2015,2016), 
                   Sales=c(792182425,915747082,919598273))

df_4 <- rbind(df_2, df_3)
df_4
```

##4. Split a string, then concatenate the results back together.
```{r}
df %>%
  separate(addr_full, sep = ' ', into = c('House Number', 'Street Name')) %>%
  mutate(Address = paste(`House Number`, `Street Name`, sep = '-'))
```
```{r}
df %>%
  separate(`Sale Date`, sep = '-', into = c('Year', 'Month', 'Day')) %>%
  mutate(`Sale Transaction Date` = paste(`Month`, `Day`, `Year`, sep = '/'))
```

