---
title: "Assignment06"
author: "Melissa Young"
date: "July 26, 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set the working directory to the root of your DSC 520 directory
setwd("/home/jdoe/Workspaces/dsc520")
```{r}
setwd('C:/Users/mdyoung/OneDrive - Bankers Financial Corporation/Documents/R/DSC 520')
```

## Load the `data/r4ds/heights.csv` to
heights_df <- read.csv("data/r4ds/heights.csv")
```{r}
heights_df <- read.csv('heights.csv')
```

## Load the ggplot2 library
library(ggplot2)
```{r}
library(ggplot2)
```

## Fit a linear model using the `age` variable as the predictor and `earn` as the outcome
age_lm <-  ___
```{r}
age_lm <- lm(`earn` ~ `age`, data = heights_df)
```

## View the summary of your model using `summary()`
___
```{r}
summary(age_lm)
```

## Creating predictions using `predict()`
age_predict_df <- data.frame(earn = predict(___, ___), age=___)
```{r}
age_predict_df <- data.frame(earn = predict(age_lm, heights_df), age=heights_df$age)
```

## Plot the predictions against the original data
ggplot(data = ___, aes(y = ___, x = ___)) +
  geom_point(color='blue') +
  geom_line(color='red',data = ___, aes(y=___, x=___))

mean_earn <- mean(heights_df$earn)
```{r}
ggplot(data = heights_df, aes(y = `earn`, x = `age`)) +
  geom_point(color='blue') +
  geom_line(color='red',data = age_predict_df, aes(y=`earn`, x=`age`))

mean_earn <- mean(heights_df$earn)
```

## Corrected Sum of Squares Total
sst <- sum((mean_earn - heights_df$earn)^2)
```{r}
sst <- sum((mean_earn - heights_df$earn)^2)
```

## Corrected Sum of Squares for Model
ssm <- sum((mean_earn - age_predict_df$earn)^2)
```{r}
ssm <- sum((mean_earn - age_predict_df$earn)^2)
```

## Residuals
residuals <- heights_df$earn - age_predict_df$earn
```{r}
residuals <- heights_df$earn - age_predict_df$earn
```

## Sum of Squares for Error
sse <- sum(residuals^2)
```{r}
sse <- sum(residuals^2)
```


r_squared <- ___
```{r}
r_squared <- ssm/sst

```

## Number of observations
n <- ___
```{r}
n <- nrow(heights_df) * ncol(heights_df)
```

## Number of regression parameters
p <- 2
```{r}
p <- 2
```

## Corrected Degrees of Freedom for Model (p-1)
dfm <- ___
```{r}
dfm <- p - 1
```

## Degrees of Freedom for Error (n-p)
dfe <- ___
```{r}
dfe <- n - p
```

## Corrected Degrees of Freedom Total:   DFT = n - 1
dft <- ___
```{r}
dft <- n - 1
```

## Mean of Squares for Model:   MSM = SSM / DFM
msm <- ___
```{r}
msm <- ssm/dfm
```

## Mean of Squares for Error:   MSE = SSE / DFE
mse <- ___
```{r}
mse <- sse/dfe
```

## Mean of Squares Total:   MST = SST / DFT
mst <- ___
```{r}
mst <- sst/dft
```

## F Statistic F = MSM/MSE
f_score <- ___
```{r}
f_score <- msm/mse
```

## Adjusted R Squared R2 = 1 - (1 - R2)(n - 1) / (n - p)
adjusted_r_squared <- ___
```{r}
adjusted_r_squared <- 1 - (1 - r_squared) * (dft)/(dfe)
```

## Calculate the p-value from the F distribution
p_value <- pf(f_score, dfm, dft, lower.tail=F)
```{r}
p_value <- pf(f_score, dfm, dft, lower.tail=F)
```

