---
title: "2014 American Community Survey_YoungMelissa.Rmd"
author: "Melissa Young"
date: "June 24, 2022"
output: pdf_document
---

## Load the packages
library(ggplot2)
library(pastecs)
library(psych)

```{r}
library(ggplot2)
library(pastecs)
library(psych)
library(formatR)
```

```{r, echo = F}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

## Load the csv fileto
df <- read.csv("acs-14-1yr-s0201.csv")
```{r}
df <- read.csv("acs-14-1yr-s0201.csv")
```


##1. What are the elements in your data (including the categories and data types)?

```{r}
summary(df)
```

##2. Please provide the output from the following functions: str(); nrow(); ncol()
```{r}
str(df)
nrow(df)
ncol(df)
```

##3. Create a Histogram of the HSDegree variable using the ggplot2 package.\
3.1 Set a bin size for the Histogram.\
3.2 Include a Title and appropriate X/Y axis labels on your Histogram Plot.
```{r}
ggplot(data = df, aes (x=HSDegree)) + geom_histogram(bins = 50, color = "black", fill = "blue", alpha = 0.7) + ggtitle("Histogram of HSDegree") + xlab("Percent of Population with HSDegree") + ylab("Count")
```

##4. Answer the following questions based on the Histogram produced: \

##4.1 Based on what you see in this histogram, is the data distribution unimodal? \
Answer - Yes, there is only one hump. \
##4.2 Is it approximately symmetrical? \
Answer - No, it seems to be negatively skewed to the left. \
##4.3 Is it approximately bell-shaped? \
Answer - No \
##4.4 Is it approximately normal? \
Answer - No, would be more bell shaped and symmetrical. \
##4.5 If not normal, is the distribution skewed? If so, in which direction? \
Answer - It's negatively skewed to the left. \ 
##4.6 Include a normal curve to the Histogram that you plotted. \

```{r}
ggplot(data = df, aes(x = HSDegree)) + geom_histogram(aes(y = ..density..), bins = 50, colour = "black", fill = "blue", alpha = 0.7) +
  stat_function(fun = dnorm, args = list(mean = mean(df$HSDegree), sd = sd(df$HSDegree)), color = "red") + 
  ggtitle("Histogram of HSDegree") +
  xlab ("Percent of Population with HSDegree") +
  ylab ("Count")
```


##4.7 Explain whether a normal distribution can accurately be used as a model for this data. \
Answer - A normal distribution would not work with this dataset, as it is skewed.\

##5. Create a Probability Plot of the HSDegree variable.
```{r}
ggplot(data = df, aes(x = HSDegree)) + geom_density(fill = "Orange") + ggtitle("Probabilty Plot of HSDegree") + xlab("Percent of Population with HSDegree") + ylab("Density") 
```

##6. Answer the following questions based on the Probability Plot: \

##6.1 Based on what you see in this probability plot, is the distribution approximately normal? Explain how you know.\
Answer - No, this plot is not normal because it is not symmetrical. There is a tail from the left. \
##6.2 If not normal, is the distribution skewed? If so, in which direction? Explain how you know. \
Answer - Yes, skewed to the left. This plot is negative skew, with the longer tail on the left of the distribution. \

##7. Now that you have looked at this data visually for normality, you will now quantify normality with numbers using the stat.desc() function. Include a screen capture of the results produced.
```{r}
round(stats <- stat.desc(df$HSDegree, basic = FALSE, norm = TRUE), digits = 3)
```

z_score <- round((df$HSDegree-mean(df$HSDegree))/sd(df$HSDegree), digits = 3)
```{r}
z_score <- round((df$HSDegree-mean(df$HSDegree))/sd(df$HSDegree), digits = 3)
z_score
```



##8. In several sentences provide an explanation of the result produced for skew, kurtosis, and z-scores. In addition, explain how a change in the sample size may change your explanation? \
Answer - For the Skew we see a value of -1.675 this indicates that the distribution is highly negatively skewed to the left.\
The Kurtosis is 4.353, this indicates a large tail with outliers. \
The z-scores show high levels of variability with the outliers. \
If you change to sample size, the summary stastics can be impacted by significant outliers and the denominator changes etc. 

