---
title: "Assignment_05_YoungMelissa.Rmd"
author: "Melissa Young"
date: "July 17, 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Set the working directory to the root of your DSC 520 directory
setwd("C:/Users/mdyoung/OneDrive - Bankers Financial Corporation/Documents/R/DSC 520")
```{r}
setwd("C:/Users/mdyoung/OneDrive - Bankers Financial Corporation/Documents/R/DSC 520")
```


## Load the `data/r4ds/heights.csv` to
heights_df <- read.csv("data/r4ds/heights.csv")
```{r}
heights_df <- read.csv("heights.csv")
```

```{r}
head(heights_df,10)
```


## Using `cor()` compute correclation coefficients for
## height vs. earn
___
```{r}
cor(x = heights_df$height, y = heights_df$earn)
```

### age vs. earn
___
```{r}
cor(x = heights_df$age, y = heights_df$earn)
```

### ed vs. earn
___
```{r}
cor(x = heights_df$ed, y = heights_df$earn)
```

## Spurious correlation
## The following is data on US spending on science, space, and technology in millions of today's dollars
## and Suicides by hanging strangulation and suffocation for the years 1999 to 2009
## Compute the correlation between these variables
tech_spending <- c(18079, 18594, 19753, 20734, 20831, 23029, 23597, 23584, 25525, 27731, 29449)
suicides <- c(5427, 5688, 6198, 6462, 6635, 7336, 7248, 7491, 8161, 8578, 9000)
___
```{r}
tech_spending <- c(18079, 18594, 19753, 20734, 20831, 23029, 23597, 23584, 25525, 27731, 29449)
suicides <- c(5427, 5688, 6198, 6462, 6635, 7336, 7248, 7491, 8161, 8578, 9000)
cor(tech_spending, suicides)
```


