---
title: "Assignment07"
author: "Melissa Young"
date: "July 27, 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Set the working directory to the root of your DSC 520 directory
setwd("/homoe/jdoe/Workspaces/dsc520")
```{r}
setwd('C:/Users/mdyoung/OneDrive - Bankers Financial Corporation/Documents/R/DSC 520')
```

## Load the `data/r4ds/heights.csv` to
heights_df <- read.csv("data/r4ds/heights.csv")
```{r}
heights_df <- read.csv('heights.csv')
```


# Fit a linear model
earn_lm <-  lm(___ ~ ___ + ___ + ___ + ___ + ___, data=___)
```{r}
earn_lm <-  lm(earn ~ ed + race + height + age + sex, data=heights_df)
```

# View the summary of your model
summary(earn_lm)

predicted_df <- data.frame(
  earn = predict(___, ___),
  ed=___, race=___, height=___,
  age=___, sex=___
  )
  
```{r}
summary(earn_lm)

predicted_df <- data.frame(
  earn = predict(earn_lm, heights_df),
  ed=heights_df$ed, race=heights_df$race, height=heights_df$height,
  age=heights_df$age, sex=heights_df$sex
  )
```

## Compute deviation (i.e. residuals)
mean_earn <- ___
```{r}
mean_earn <- mean(heights_df$earn)
```

## Corrected Sum of Squares Total
sst <- ___
```{r}
sst <- sum((mean_earn - heights_df$earn)^2)
```

## Corrected Sum of Squares for Model
ssm <- ___
```{r}
ssm <- sum((mean_earn - predicted_df$earn)^2)
```

## Residuals
residuals <- ___
```{r}
residuals <- heights_df$earn - predicted_df$earn
```

## Sum of Squares for Error
sse <- ___
```{r}
sse <- sum(residuals^2)
```

## R Squared
r_squared <- ___
```{r}
r_squared <- ssm/sst
```

## Number of observations
n <- ___
```{r}
n <- nrow(heights_df) * ncol(heights_df)
```

## Number of regression paramaters
p <- 8
```{r}
p <- 8
```

## Corrected Degrees of Freedom for Model
dfm <- ___
```{r}
dfm <- p - 1
```

## Degrees of Freedom for Error
dfe <- ___
```{r}
dfe <- n - p
```

## Corrected Degrees of Freedom Total:   DFT = n - 1
dft <- ___
```{r}
dft <- n - 1
```

## Mean of Squares for Model:   MSM = SSM / DFM
msm <- ___
```{r}
msm <- ssm/dfm
```

## Mean of Squares for Error:   MSE = SSE / DFE
mse <- ___
```{r}
mse <- sse/dfe
```

## Mean of Squares Total:   MST = SST / DFT
mst <- ___
```{r}
mst <- sst/dft
```

## F Statistic
f_score <- ___
```{r}
f_score <- msm/mse
```

## Adjusted R Squared R2 = 1 - (1 - R2)(n - 1) / (n - p)
adjusted_r_squared <- ___
```{r}
adjusted_r_squared <- 1 - (1 - r_squared) * (dft)/(dfe)
```

