---
title: "Assignment_04_YoungMelissa"
author: "Melissa Young"
date: "July 11, 2022"
output: pdf_document
---

```{r, echo = F}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

```{r,warning=FALSE,message=FALSE}
library(ggplot2)
```

## Load the ggplot2 package
library(ggplot2)
theme_set(theme_minimal())
```{r}
theme_set(theme_minimal())
```


## Set the working directory to the root of your DSC 520 directory
setwd("/home/jdoe/Workspaces/dsc520")
```{r}
setwd("C:/Users/mdyoung/OneDrive - Bankers Financial Corporation/Documents/R/DSC 520")
```


## Load the `data/r4ds/heights.csv` to
heights_df <- read.csv("heights.csv")
```{r}
heights_df <- read.csv("heights.csv")
```


# https://ggplot2.tidyverse.org/reference/geom_boxplot.html
## Create boxplots of sex vs. earn and race vs. earn using `geom_point()` and `geom_boxplot()`
## sex vs. earn
ggplot(___, aes(x=___, y=___)) + ___+ ___
```{r}
ggplot(heights_df, aes(x=sex, y=earn)) + geom_point()+ geom_boxplot()
```

## race vs. earn
ggplot(___, aes(x=___, y=___)) + ___+ ___
```{r}
ggplot(heights_df, aes(x=race, y=earn)) + geom_point()+ geom_boxplot()
```


# https://ggplot2.tidyverse.org/reference/geom_bar.html
## Using `geom_bar()` plot a bar chart of the number of records for each `sex`
ggplot(___, aes(___)) + ___
```{r}
ggplot(heights_df, aes(sex)) + geom_bar()
```


## Using `geom_bar()` plot a bar chart of the number of records for each race
ggplot(heights_df, aes(___)) + ___
```{r}
ggplot(heights_df, aes(race)) + geom_bar()
```


## Create a horizontal bar chart by adding `coord_flip()` to the previous plot
ggplot(___, aes(___)) + ___ + ___
```{r}
ggplot(heights_df, aes(race)) + geom_bar() + coord_flip()
```


# https://www.rdocumentation.org/packages/ggplot2/versions/3.3.0/topics/geom_path
## Load the file `"data/nytimes/covid-19-data/us-states.csv"` and
## assign it to the `covid_df` dataframe
covid_df <- read.csv("data/nytimes/covid-19-data/us-states.csv")
```{r}
covid_df <- read.csv("data/nytimes/covid-19-data/us-states.csv")
head(covid_df, 10)
```


## Parse the date column using `as.Date()``
covid_df$date <- ___
```{r}
covid_df$date <- as.Date(x = covid_df$date, "%Y-%m-%d")
```


## Create three dataframes named `california_df`, `ny_df`, and `florida_df`
## containing the data from California, New York, and Florida
california_df <- covid_df[ which( covid_df$state == "California"), ]
ny_df <- ___
florida_df <- ___
```{r}
california_df <- covid_df[ which( covid_df$state == "California"), ]
ny_df <- covid_df[ which( covid_df$state == "New York"), ]
florida_df <- covid_df[ which( covid_df$state == "Florida"), ]
```


## Plot the number of cases in Florida using `geom_line()`
ggplot(data=florida_df, aes(x=___, y=___, group=1)) + ___
```{r}
ggplot(data=florida_df, aes(x=date, y=cases, group=1)) + geom_line()
```


## Add lines for New York and California to the plot
ggplot(data=___, aes(x=___, group=1)) +
  geom_line(aes(y = ___)) +
  geom_line(data=___, aes(y = ___)) +
  geom_line(data=___, aes(y = ___))
```{r}
ggplot(data=florida_df, aes(x=date, group=1)) +
  geom_line(aes(y = cases)) +
  geom_line(data=ny_df, aes(y = cases)) +
  geom_line(data=california_df, aes(y = cases))
```


## Use the colors "darkred", "darkgreen", and "steelblue" for Florida, New York, and California
ggplot(data=___, aes(x=___, group=1)) +
  geom_line(aes(y = ___), color = ___) +
  geom_line(data=ny_df, aes(y = cases), color=___) +
  geom_line(data=___, aes(y = ___), color=___)
```{r}
ggplot(data=florida_df, aes(x=date, group=1)) +
  geom_line(aes(y = cases), color = "darkred") +
  geom_line(data=ny_df, aes(y = cases), color = "darkgreen") +
  geom_line(data=california_df, aes(y = cases), color = "steelblue")
```


## Add a legend to the plot using `scale_colour_manual`
## Add a blank (" ") label to the x-axis and the label "Cases" to the y axis
ggplot(data=___, aes(x=___, group=1)) +
  geom_line(aes(y = cases, colour = "Florida")) +
  geom_line(data=ny_df, aes(y = cases,colour="New York")) +
  geom_line(data=california_df, aes(y = cases, colour="California")) +
  scale_colour_manual("",
                      breaks = c(___, ___, ___),
                      values = c(___, ___, ___)) +
  xlab(" ") + ylab("Cases")
```{r}
ggplot(data=florida_df, aes(x=date, group=1)) +
  geom_line(aes(y = cases, colour = "Florida")) +
  geom_line(data=ny_df, aes(y = cases,colour="New York")) +
  geom_line(data=california_df, aes(y = cases, colour="California")) +
  scale_colour_manual("",
                      breaks = c("Florida", "New York", "California"),
                      values = c("darkred", "darkgreen", "steelblue")) +
  xlab(" ") + ylab("Cases")
```


## Scale the y axis using `scale_y_log10()`
ggplot(data=___, aes(x=___, group=1)) +
  geom_line(aes(y = cases, colour = "Florida")) +
  geom_line(data=ny_df, aes(y = cases,colour="New York")) +
  geom_line(data=california_df, aes(y = cases, colour="California")) +
  scale_colour_manual("",
                      breaks = c(___, ___, ___),
                      values = c(___, ___, ___)) +
  xlab(" ") + ylab("Cases") + ___
```{r}
ggplot(data=florida_df, aes(x=date, group=1)) +
  geom_line(aes(y = cases, colour = "Florida")) +
  geom_line(data=ny_df, aes(y = cases,colour="New York")) +
  geom_line(data=california_df, aes(y = cases, colour="California")) +
  scale_colour_manual("",
                      breaks = c("Florida", "New York", "California"),
                      values = c("darkred", "darkgreen", "steelblue")) +
  xlab(" ") + ylab("Cases") + scale_y_log10()
```
  
  