---
title: "Assignment00_YoungMelissa"
author: "Melissa Young"
date: "June 13, 2022"
output: pdf_document
---
Basics

## Add 8 and 5
```{r}
8 + 5
```

## Subtract 6 from 22
```{r}
6 - 22
```

## Multiply 6 by 7
```{r}
6 * 7
```


## Add 4 to 6 and divide the result by 2
```{r}
(4 + 6) / 2
```


## Compute 5 modulo 2
```{r}
5 %% 2
```


## Assign the value 82 to the variable x
## Print x
```{r}
x <- 82
x
```


## Assign the value 41 to the variable y
## Print y
```{r}
y <- 41
y
```


## Assign the output of x + y to the variable z
## Print z
```{r}
z <- x + y
z
```


## Assign the string value "DSC520" to the variable class_name
## Print the value of class_name
```{r}
class_name <- 'DSC520'
class_name
```


## Assign the string value of TRUE to the variable is_good
## Print the value of is_good
```{r}
is_good <- TRUE
is_good
```


## Check the class of the variable is_good using the `class()` function
```{r}
class (is_good)
```


## Check the class of the variable z using the `class()` function
```{r}
class (z)
```


## Check the class of the variable class_name using the class() function
```{r}
class(class_name) 
```

